### 2025-09-13 paper/sec/evaluation.tex
[draft] added validation paragraph

@@

\subsection{Success Rate by Category}
To evaluate the system we used a sample of the Scigen dataset \cite{scigen_dataset_2021}.

@@

>>>>

\subsection{Validation}
\label{subsec:validation-with-counterfactual-data}

The validation process is structured in two phases

\paragraph{Initial validation}
In the initial phase, the generated expression is executed via Fluid CLI.
If the execution is successful, the result obtained is compared with the expected result.
In case of error, the output is forwarded to the interpretation agent, which proceeds to regenerate the expression.

\paragraph{Validation via counterfactual testing}
In the second phase, the data used for execution is deliberately modified in order to verify the consistency of the behaviours.
Both the expected and generated expressions are executed again and the results are compared.
If discrepancies emerge, the error is forwarded to the interpretation agent.
If there are no differences, the expression is considered valid and is returned to the user.

@@

\subsection{Success Rate by Category}
To evaluate the system we used a sample of the Scigen dataset~\citep{scigen_dataset_2021}.

@@

\subsection{Negative Cases}
\label{subsec:negative-cases}

\begin{table}[t]
    \centering
    \small
    \begin{tabular}{p{3cm} p{6cm} p{4cm}}
        \toprule
        \textbf{Ambiguity Type} & \textbf{Example} & \textbf{Explanation} \\
        \midrule

        ambiguous referent &
        As shown in Table 3 \hl{BiLSTM} gives significantly better accuracies compared to uni-directional LSTM with the training time per epoch growing from 99 seconds to 106 seconds &
        There are three models that have 99 seconds as their time; the assistant would not know which one to choose. \\

        \midrule
        missing-data &
        The average methane emissions for the year 2015 is \hl{1.0} &
        November is missing to calculate the average correctly. \\

        \midrule
        false statement &
        We additionally make comparisons with stacked CNNs and hierarchical attention (Vaswani et al., 2017), shown in Table 3 (the CNN and Transformer rows), BiLSTM is \hl{the most efficient} among all models compared, with the highest model size &
        BiLSTM is not the most efficient, nor does it have the largest size. \\

        \midrule
        incorrect numerical value &
        LSTM is the fastest model with overall time taken being \hl{90} seconds &
        It is not 90 but 106. \\

        \midrule
        ambiguous referent &
        LSTM is the fastest model with overall time taken being \hl{90} seconds &
        There are two type of time in the dataset (training\_time, execution\_time), both with a value of 90 seconds. \\

        \bottomrule
    \end{tabular}
    \caption{Negative cases example without the \emph{Data} column.}
    \label{tab:ambiguita}
\end{table}
